target_offtarget

### 2025-01-03 manuscript_figures.Rmd
update manuscript_figures.Rmd

counts_reads_norm <- counts_reads |>
  mutate(norm_counts1 = matched / QC_reads) |>
  mutate(norm_counts2 = matched / QC_reads / length) |>
  mutate(norm_counts3 = matched / length) |>
  mutate(genome_structure = case_when((
    virus == "Human_adenovirus_40" |
      virus == "Human_betaherpesvirus"
  ) ~ "DNA",
  ,
  .default = "RNA"
  ))

@@

## Figure S2
```{r}
#metadata

metadata <-
  read.csv(
    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/metadata/sampleIDs_TESpikeIn.csv",
    header = TRUE
  )

metadata2 <- metadata %>%
  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) %>%
  rename(Sample_id = Sample.ID) %>%
  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)

@@

counts_dedup_bt <-
  read.table(
    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

counts_nodedup_bt <-
  read.table(
    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

@@

counts_reads_norm <- counts_reads %>%
  mutate(norm_counts1 = matched / QC_reads) %>%
  mutate(norm_counts2 = matched / QC_reads / length) %>%
  mutate(norm_counts3 = matched / length) %>%
  mutate(genome_structure = case_when((
    virus == "Human_adenovirus_40" |
      virus == "Human_betaherpesvirus"
  ) ~ "DNA",
  ,
  .default = "RNA"
  ))

@@

metadata3 <- metadata %>%
  select(Sample.ID, Pool.for.sequencing) %>%
  rename(Sample_id = Sample.ID) %>%
  rename(Pool = Pool.for.sequencing)

@@

pool_readcounts <- counts_pool %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(x = virus, y = log(matched))) +
  geom_boxplot() +
  facet_grid(Pool ~ type) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_text(size = 10)
  ) +
  ylab("Log(Viral Reads)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))

pool_counts_sum <- counts_pool %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(
    x = as.character(Viral.load),
    y = log(matched),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_bw() +
  theme(axis.title.x = element_blank(), axis.title.y = element_text(size =
                                                                      10)) +
  ylab("Log(Viral Reads)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))

pool_readcounts_norm <- counts_pool %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(x = virus, y = log(norm_counts1))) +
  geom_boxplot() +
  facet_grid(Pool ~ type) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_text(size = 10)
  ) +
  ylab("Log(viral reads/cleaned reads)")

pool_norm1_sum <- counts_pool %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(
    x = as.character(Viral.load),
    y = log(matched),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_bw() +
  theme(axis.title.x = element_blank(), axis.title.y = element_text(size =
                                                                      10)) +
  ylab("Log(viral reads/cleaned reads)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))

pool_readcounts_norm2 <- counts_pool %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(x = virus, y = log(norm_counts2))) +
  geom_boxplot() +
  facet_grid(Pool ~ type) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10)
  ) +
  ylab("Log(viral reads/cleaned reads/genome length)")

pool_norm2_sum <- counts_pool %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(
    x = as.character(Viral.load),
    y = log(matched),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_bw() +
  theme(axis.title.x = element_blank(), axis.title.y = element_text(size =
                                                                      10)) +
  ylab("Log(viral reads/cleaned reads/genome length)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))

@@

depth_dedup_bt <-
  read.table(
    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

depth_nodedup_bt <-
  read.table(
    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

@@

depths_reads <- left_join(depths_bt_all, metadata2, by = "Sample_id") %>%
  mutate(genome_structure = case_when((
    virus == "Human_adenovirus_40" |
      virus == "Human_betaherpesvirus"
  ) ~ "DNA",
  ,
  .default = "RNA"
  ))

@@

pool_readdepths <- depths_pool %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(x = virus, y = log(mean_depth))) +
  geom_boxplot() +
  facet_grid(Pool ~ type) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10)
  ) +
  ylab("Log(mean read depth)")

pool_depths_sum <- depths_pool %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(
    x = as.character(Viral.load),
    y = log(mean_depth),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_bw() +
  theme(axis.title.x = element_blank(), axis.title.y = element_text(size =
                                                                      10)) +
  ylab("Log(mean read depth)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))

ggarrange(
  pool_readcounts,
  pool_readcounts_norm,
  pool_readcounts_norm2,
  pool_readdepths,
  nrow = 2,
  ncol = 2,
  heights = c(1, 2),
  common.legend = TRUE
)

#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/compare_spike_ins_atcc/pool_compare_viruses.pdf")

ggarrange(
  pool_counts_sum,
  pool_norm1_sum,
  pool_norm2_sum,
  pool_depths_sum,
  nrow = 2,
  ncol = 2,
  common.legend = TRUE
)

#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/compare_spike_ins_atcc/pool_compare_viruses_sum.png",width=10,height=7)

#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/manuscript_figures_pdf/FigureS2.pdf",width=10,height=7)

@@

library(tidyverse)
library(ggpubr)

@@

metadata <-
  read.csv(
    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/metadata/sampleIDs_TESpikeIn.csv",
    header = TRUE
  )

metadata2 <- metadata %>%
  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) %>%
  rename(Sample_id = Sample.ID) %>%
  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)

@@

viral_reads_dedup <-
  read.csv(
    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/total_virus_mapped_reads_per_sample_dedup_atcc_ref_20241108.csv",
    header = TRUE
  )

viral_reads_nodedup <-
  read.csv(
    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/total_virus_mapped_reads_per_sample_nodedup_atcc_ref_20241108.csv",
    header = TRUE
  )

@@

reads_plot <- reads_viral_all %>%
  group_by(Background, Sample_id, Viral.load, type) %>%
  summarise(
    total_reads = (QC_reads * 2),
    viral_reads = total_virus_reads,
    ATCC_reads = sum(matched),
    prop_ATCC = ATCC_reads / total_reads,
    prop_viral = total_virus_reads / total_reads,
    diff = viral_reads - ATCC_reads
  ) %>%
  unique()

total_reads <- reads_plot %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(
    x = Viral.load,
    y = log(total_reads),
    colour = Background
  )) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10)
  ) +
  scale_x_log10() +
  ggtitle("Total Reads") +
  ylab("Log(Total Reads)") +
  scale_color_manual(values = cols2, labels = c("ME_P1", "ME_P2"))

viral_reads <- reads_plot %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(
    x = Viral.load,
    y = log(viral_reads),
    colour = Background
  )) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  theme_bw() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10)
  ) +
  scale_x_log10() +
  ggtitle("Total Viral Reads") +
  ylab("Log(Total Viral Reads)") +
  scale_color_manual(values = cols2, labels = c("ME_P1", "ME_P2"))

ATCC_reads <- reads_plot %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(
    x = Viral.load,
    y = log(ATCC_reads),
    colour = Background
  )) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  theme_bw() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10)
  ) +
  scale_x_log10() +
  ggtitle("Spike In Viral Reads") +
  ylab("Log(Spike In Viral Reads)") +
  scale_color_manual(values = cols2, labels = c("ME_P1", "ME_P2"))

prop_ATCC <- reads_plot %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(x = Viral.load, y = prop_ATCC, colour = Background)) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10)
  ) +
  scale_x_log10() +
  ggtitle("Proportion Spike In Viral") +
  ylab("Spike In Viral Reads/Total Reads") +
  scale_color_manual(values = cols2, labels = c("ME_P1", "ME_P2"))

prop_viral <- reads_plot %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(x = Viral.load, y = prop_viral, colour = Background)) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  theme_bw() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10)
  ) +
  scale_x_log10() +
  ggtitle("Proportion Viral") +
  ylab("Viral Reads/Total Reads") +
  scale_color_manual(values = cols2, labels = c("ME_P1", "ME_P2"))

@@

ggarrange(total_reads,
          viral_reads,
          prop_viral,
          nrow = 3,
          common.legend = TRUE)

#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/compare_spike_ins_atcc/backgrounds_reads.png")

#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/manuscript_figures_pdf/FigureS3.pdf")

@@

metadata3 <- metadata %>%
  select(Sample.ID,
         Number.of.read.pairs..quality.adaptor.trimmed.,
         Pool.for.sequencing) %>%
  rename(Sample_id = Sample.ID) %>%
  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)

reads_viral_dedup3 <-
  full_join(metadata3, viral_reads_dedup, by = "Sample_id")

reads_viral_nodedup3 <-
  full_join(metadata3, viral_reads_nodedup, by = "Sample_id")

reads_viral_dedup3 %>%
  group_by(Pool.for.sequencing) %>%
  summarise(
    viral_reads = sum(total_virus_reads),
    total_reads = sum(QC_reads * 2),
    prop_viral = viral_reads / total_reads
  )

polyomics <-
  read.csv(
    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_polyomics/total_virus_mapped_reads_per_sample_dedup.csv"
  )

#read data from polyomics_indexes_stefano document
total_reads <- data.frame(total_reads = c(5942760, 7714275))

keeps <- c("RNA-Msp-p2", "RNA-Msp-p8")

polyomics_read_prop <- polyomics %>%
  filter(Sample_id %in% keeps)

polyomics_read_prop2 <- cbind(polyomics_read_prop, total_reads)

polyomics_read_prop2 %>%
  summarise(prop_viral = total_virus_reads / total_reads)

@@

background_counts_reads <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(matched),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid(Background ~ type, labeller = as_labeller(facet_names_bg)) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10),
    legend.title = element_blank()
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  scale_x_log10() +
  ylab("Log(Viral Reads)")

backgrounds_counts_reads_norm <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(norm_counts1),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid(Background ~ type, labeller = as_labeller(facet_names_bg)) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10),
    legend.title = element_blank()
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  scale_x_log10() +
  ylab("Log(viral reads/cleaned reads)")

backgrounds_counts_reads_norm2 <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(norm_counts2),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid(Background ~ type, labeller = as_labeller(facet_names_bg)) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10),
    legend.title = element_blank()
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  scale_x_log10() +
  ylab("Log(viral reads/cleaned reads/genome length)")

>>>>

@@

## Figure S2

### Reaed-in data / Make plots
```{r}
#metadata

@@

@@

@@

metadata3 <- 
  metadata |>
  select(Sample.ID, Pool.for.sequencing) |>
  rename(Sample_id = Sample.ID) |>
  rename(Pool = Pool.for.sequencing)

@@

pool_readcounts <- 
  counts_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x = virus, y = log(matched))) +
  geom_boxplot() +
  facet_grid(Pool ~ type) +
  theme_few() +
  theme(
    # axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("Log (Viral Reads)") + xlab("Viral load (copies / ml)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))

pool_counts_sum <- 
  counts_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = as.character(Viral.load),
    y = log(matched),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    # axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("Log (Viral Reads)") + xlab("Viral load (copies / ml)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))

pool_readcounts_norm <- 
  counts_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x = virus, y = log(norm_counts1))) +
  geom_boxplot() +
  facet_grid(Pool ~ type) +
  theme_few() +
  theme(
    # axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("Log(viral reads/cleaned reads)") +  xlab("Viral load (copies / ml)")

pool_norm1_sum <- 
  counts_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = as.character(Viral.load),
    y = log(matched),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    # axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("Log(viral reads/cleaned reads)") +   xlab("Viral load (copies / ml)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))

pool_readcounts_norm2 <- 
  counts_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x = virus, y = log(norm_counts2))) +
  geom_boxplot() +
  facet_grid(Pool ~ type) +
  theme_few() +
  theme(
    # axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("Log(viral reads/cleaned reads/genome length)") + xlab("Viral load (copies / ml)")

pool_norm2_sum <- 
  counts_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = as.character(Viral.load),
    y = log(matched),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    # axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("Log(viral reads/cleaned reads/genome length)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2")) + 
  xlab("Viral load (copies / ml)")

@@

depth_dedup_bt <-
  read.table(
    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

depth_nodedup_bt <-
  read.table(
    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

@@

depths_reads <- 
  left_join(depths_bt_all, metadata2, by = "Sample_id") |>
  mutate(genome_structure = case_when((
    virus == "Human_adenovirus_40" |
      virus == "Human_betaherpesvirus"
  ) ~ "DNA",
  .default = "RNA"
  ))

@@

pool_readdepths <- 
  depths_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x = virus, y = log(mean_depth))) +
  geom_boxplot() +
  facet_grid(Pool ~ type) +
  theme_few() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("Log (mean read depth)") + xlab("Viral load (copies / ml)")

pool_depths_sum <- 
  depths_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = as.character(Viral.load),
    y = log(mean_depth),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    # axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("Log (mean read depth)") +
  xlab("Viral load (copies / ml)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))

### Final plot
```{r}
ggarrange(
  pool_counts_sum,
  pool_norm1_sum,
  pool_norm2_sum,
  pool_depths_sum,
  nrow = 2,
  ncol = 2,
  common.legend = TRUE,
  align = "hv"
)

#ggsave("figures/compare_spike_ins_atcc/pool_compare_viruses_sum.png",width=10,height=7)

#ggsave("figures/manuscript_figures_pdf/FigureS2.pdf",width=10,height=7)

@@

@@

@@

@@

reads_plot <- 
  reads_viral_all |>
  group_by(Background, Sample_id, Viral.load, type) |>
  summarise(
    total_reads = (QC_reads * 2),
    viral_reads = total_virus_reads,
    ATCC_reads = sum(matched),
    prop_ATCC = ATCC_reads / total_reads,
    prop_viral = total_virus_reads / total_reads,
    diff = viral_reads - ATCC_reads
  ) |>
  unique()

### Make plots
```{r, echo=FALSE, warning=FALSE, message=FALSE}

total_reads <- 
  reads_plot |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(total_reads),
    colour = Background
  )) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10(labels = scales::label_log()) + xlab("Viral load (copies / ml)") +
  ggtitle("Total Reads") +
  ylab("Log(Total Reads)") +
  scale_color_manual(values = cols2, labels = c("P1", "P2"))

viral_reads <- 
  reads_plot |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(viral_reads),
    colour = Background
  )) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10(labels = scales::label_log()) + xlab("Viral load (copies / ml)") +
  ggtitle("Total Viral Reads") +
  ylab("Log(Total Viral Reads)") +
  scale_color_manual(values = cols2, labels = c("P1", "P2"))

ATCC_reads <- 
  reads_plot |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(ATCC_reads),
    colour = Background
  )) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10(labels = scales::label_log()) + xlab("Viral load (copies / ml)") +
  ggtitle("Spike In Viral Reads") +
  ylab("Log(Spike In Viral Reads)") +
  scale_color_manual(values = cols2, labels = c("P1", "P2"))

prop_ATCC <- reads_plot |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x = Viral.load, y = prop_ATCC, colour = Background)) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10(labels = scales::label_log()) + xlab("Viral load (copies / ml)") +
  ggtitle("Proportion Spike In Viral") +
  ylab("Spike In Viral Reads/Total Reads") +
  scale_color_manual(values = cols2, labels = c("P1", "P2"))

prop_viral <- reads_plot |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x = Viral.load, y = prop_viral, colour = Background)) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10(labels = scales::label_log()) + xlab("Viral load (copies / ml)") +
  ggtitle("Proportion Viral") +
  ylab("Viral Reads/Total Reads") +
  scale_color_manual(values = cols2, labels = c("P1", "P2"))

@@

### ***Final plot***
```{r}
ggarrange(total_reads,
          viral_reads,
          prop_viral,
          nrow = 3,
          common.legend = TRUE,
          align = "hv")

#ggsave("figures/compare_spike_ins_atcc/backgrounds_reads.png")

#ggsave("figures/manuscript_figures_pdf/FigureS3.pdf")

@@

# metadata3 <- metadata |>
#   select(Sample.ID,
#          Number.of.read.pairs..quality.adaptor.trimmed.,
#          Pool.for.sequencing) |>
#   rename(Sample_id = Sample.ID) |>
#   rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)
# 
# reads_viral_dedup3 <-
#   full_join(metadata3, viral_reads_dedup, by = "Sample_id")
# 
# reads_viral_nodedup3 <-
#   full_join(metadata3, viral_reads_nodedup, by = "Sample_id")
# 
# reads_viral_dedup3 |>
#   group_by(Pool.for.sequencing) |>
#   summarise(
#     viral_reads = sum(total_virus_reads),
#     total_reads = sum(QC_reads * 2),
#     prop_viral = viral_reads / total_reads
#   )
# 
# polyomics <-
#   read.csv(
#     "data_polyomics/total_virus_mapped_reads_per_sample_dedup.csv"
#   )
# 
# #read data from polyomics_indexes_stefano document
# total_reads <- data.frame(total_reads = c(5942760, 7714275))
# 
# keeps <- c("RNA-Msp-p2", "RNA-Msp-p8")
# 
# polyomics_read_prop <- polyomics |>
#   filter(Sample_id %in% keeps)
# 
# polyomics_read_prop2 <- cbind(polyomics_read_prop, total_reads)
# 
# polyomics_read_prop2 |>
#   summarise(prop_viral = total_virus_reads / total_reads)

@@

background_counts_reads <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(matched),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid(Background ~ type, labeller = as_labeller(facet_names_bg)) +
  theme_few() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels =
      c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
      )
  ) +
  scale_x_log10(labels = scales::label_log()) +
  ylab("Log(Viral Reads)")

backgrounds_counts_reads_norm <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(norm_counts1),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid(Background ~ type, labeller = as_labeller(facet_names_bg)) +
  theme_bw() +
  theme_few() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  scale_x_log10(label = scales::label_log()) +
  ylab("Log(viral reads/cleaned reads)")

backgrounds_counts_reads_norm2 <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(norm_counts2),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid(Background ~ type, labeller = as_labeller(facet_names_bg)) +
  theme_few() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  scale_x_log10(labels = scales::label_log()) +
  ylab("Log(viral reads/cleaned reads/genome length)")
